\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2019-01-23 Documents/M1/Astronum_2018/sections/Conclusion.tex

We have developed IMEX schemes suitable for a two-moment model of neutrino transport in \texttt{thornado} that can obey Fermi-Dirac statistics.  \ee{(What is \texttt{thornado}?  It is mentioned here and in the abstract, and not anywhere else.  Need a paragraph in the intro?)}
The scheme employs algebraic closure based on Fermi-Dirac statistics, high-order discontinuous Galerkin method for spatial discretization, and convex-invariant time integration to maintain realizability of the moments.  
Since the realizability-preserving property is obtained from the convexity of the realizable domain, it is possible to construct a scheme with the high-order DG method \ee{(Not sure what this means.  Does it have to be DG?)}.  
See \cite{chu_etal_2018} for some examples with third-order DG methods.

In the application motivating this work, the neutrino distribution function can vary from 0 to 1.  
Hence, we have considered algebraic closures based on Fermi-Dirac statistics for both low and high occupancy.  
Among the seven algebraic closures we considered -- Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, Janka 1~\cite{janka_1991}, Janka 2~\cite{janka_1992}, and Cernohorsky \& Bludman~\cite{cernohorskyBludman_1994} -- only the Cernohorsky \& Bludman closure obeys Fermi-Dirac statistics for all occupancies.  
As a result, we employed the Cernohorsky \& Bludman closure in all of the numerical tests in Section~\ref{se:NumericalTests}.  \ee{(But you said you used Minerbo for the accuracy tests)}

Two PD-ARS schemes are proposed.  
The one with SSPRK2 is second-order accurate and strong-stability preserving in the streaming limit while the other with SSPRK3 is third-order accurate in the streaming limit.  
Their accuracy were demonstrated on problems with known smooth solutions in streaming, absorption, and scattering-dominated regimes.
The neutrino transport test with emission, absorption, and elastic scattering through a stationary background, was designed to test the convex-invariance of our PD-ARS schemes.
The neutrino stationary state test shows that it's necessary to combine an algebraic closure based on Fermi-Dirac statistics and convex-invariant time integration for a solid CCSN simulation.  \ee{(This statement is too strong.  Rewrite.)}

>>>>

We have developed IMEX schemes suitable for a two-moment model of neutrino transport in \texttt{thornado} that can obey Fermi-Dirac statistics.
The scheme employs algebraic closure based on Fermi-Dirac statistics, high-order discontinuous Galerkin method for spatial discretization, and convex-invariant time integration to maintain realizability of the moments.  
Since the realizable domain is convex and its convexity can be inherited by a convex combination, some schemes having convex combinations as their stages should be able to preserve the convexity.
This idea encourages us to construct realizability-preserving time integrators, realizability-preserving IMEX schemes, and eventually a method with realizability-preserving IMEX time integrator and high-order DG method.
See \cite{chu_etal_2018} for some examples with third-order DG methods.

In the application motivating this work, the neutrino distribution function can vary from 0 to 1.  
Hence, we have considered algebraic closures based on Fermi-Dirac statistics for both low and high occupancy.  
Among the seven algebraic closures we considered -- Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, Janka 1~\cite{janka_1991}, Janka 2~\cite{janka_1992}, and Cernohorsky \& Bludman~\cite{cernohorskyBludman_1994} -- only the Cernohorsky \& Bludman closure obeys Fermi-Dirac statistics for all occupancies.  
As a result, we employed the Cernohorsky \& Bludman closure for the neutrino stationary state test in Section~\ref{se: Neutrino Stationary State Test}.
(We also run the code with other algebraic closures for examination.
And the program crashed after few steps because of the nonphysical neutrino density/flux.)

Two PD-ARS schemes are proposed.
The one with SSPRK2 has second-order accuracy while the other with SSPRK3 has third-order accuracy, and both have strong-stability preserving property in the streaming limit.  
Their accuracy were demonstrated on problems with known smooth solutions in streaming, absorption, and scattering-dominated regimes.
The neutrino transport test with emission, absorption, and elastic scattering through a stationary background, was designed to test the convex-invariance of our PD-ARS schemes. 
The neutrino stationary state test shows that a method combined an algebraic closure based on Fermi-Dirac statistics and convex-invariant time integration can be a candidate for constructing a CCSN simulation code without "nonphysical protection" that may break conservation laws.

### 2019-01-23 Documents/M1/Astronum_2018/sections/Numerical.tex

To compare the accuracy of the IMEX schemes, we applied our PD-ARS schemes and the SSP2332 scheme from Pareschi \& Russo \cite{pareschiRusso_2005} to problems with known smooth solutions in streaming, absorption, and scattering-dominated regimes in one spatial dimension.  
All the tests in this subsection were computed with the maximum entropy closure in the low-occupancy limit (i.e., the Minerbo closure).  
In the streaming test, the second- and third-order accurate explicit strong-stability-preserving Runge-Kutta methods from \cite{gottlieb_etal_2001} (SSPRK2 and SSPRK3, respectively) are also included.  
To compare the numerical results to analytic solutions, the cell-averaged absolute error or the cell averaged relative error are computed \ee{(give more specific details here)}.  
For our convergence test, the number of elements ($N$) vary from 8 to 128.

@@

In this section we consider a more ``realistic'' test: two-dimensional multigroup neutrino transport with emission, absorption, and elastic scattering through a stationary background.  
This test is designed to test the realizability-preserving properties of the PD-ARS schemes.  
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities used (adopted from \cite{Bruenn_1985}).  
This test is computed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction.  
\ee{(State boundary conditions.)}
We use 10 energy groups to cover the range from 0~MeV to 300~MeV and initialize the neutrino number density to $\cJ = 10^{-99}$, flux density $\bcH=0$, and evolve until an approximate steady state is reached ($t=5$~ms).  
For this test, we employed the CB closure.  
We attempted to run this test with our PD-ARS schemes, SSP2332 from \cite{pareschiRusso_2005}, IMEXRKCB2 proposed by Cavaglieri \& Bewley \cite{cavaglieriBewley2015}, and the scheme proposed by McClarren et al. \cite{mcclarren_etal_2008}.
Only the PD-ARS schemes produce realizable moments and are able to evolve to a steady state.
The other schemes crashed after a few steps, when moments became non-realizable and the closure procedure failed.

>>>>

To compare the accuracy of the IMEX schemes, we applied our PD-ARS schemes and the SSP2332 scheme from Pareschi \& Russo \cite{pareschiRusso_2005} to problems with known smooth solutions in streaming, absorption(damping), and scattering-dominated(diffusing) regimes in one spatial dimension.  
All the tests in this subsection were computed with the maximum entropy closure in the low-occupancy limit (i.e., the Minerbo closure).  
In the streaming test, the second- and third-order accurate explicit strong-stability-preserving Runge-Kutta methods from \cite{gottlieb_etal_2001} (SSPRK2 and SSPRK3, respectively) are also included.  
To compare the numerical results to analytic solutions, the averaged absolute error or the averaged relative error are computed in the $L^{1}$-error norm: the absolute error for the streaming test and the diffusion test; the relative error for the damping test.
They are averaged over the cell with a equal weight quadrature for the cell integrals.
And to examine the convergence, we let the number of elements ($N$) vary from 8 to 128.

@@

In this section we consider a more ``realistic'' test: two-dimensional multigroup neutrino transport with emission, absorption, and elastic scattering through a stationary background.  
This test is designed to test the realizability-preserving properties of the PD-ARS schemes.  
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities used (adopted from \cite{Bruenn_1985}).  
This test is computed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction, a reflecting inner boundary and a homogeneous out-flowing outer boundary.
We use 10 energy groups to cover the range from 0~MeV to 300~MeV and initialize the neutrino number density to $\cJ = 10^{-99}$, flux density $\bcH=0$, and evolve until an approximate steady state is reached ($t=5$~ms).  
For this test, we employed the CB closure.  
We attempted to run this test with our PD-ARS schemes, SSP2332 from \cite{pareschiRusso_2005}, IMEXRKCB2 proposed by Cavaglieri \& Bewley \cite{cavaglieriBewley2015}, and the scheme proposed by McClarren et al. \cite{mcclarren_etal_2008}.
Only the PD-ARS schemes produce realizable moments and are able to evolve to a steady state.
The other schemes crashed after a few steps, when moments became non-realizable and the closure procedure failed.
